---
title             : "Using contrastive inferences to learn about new words and categories"
shorttitle        : "Learning from contrastive inference"

author: 
  - name          : "----"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "---- University Avenue, ----"
    email         : "-----.edu"
  - name          : "-----"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "------"
  - id            : "2"
    institution   : "------"

abstract: |
  In the face of unfamiliar language or objects, description is one cue people can use to learn about both. Beyond narrowing potential referents to those that match a descriptor (e.g., "tall"), people could infer that a described object is one that contrasts with other relevant objects of the same type (e.g., "the tall cup" contrasts with another, shorter cup). This contrast may be in relation to other objects present in the environment (this cup is tall among present cups) or to the referent's category (this cup is tall for a cup in general). In three experiments, we investigate whether people use such contrastive inferences from description to learn new word-referent mappings and learn about new categories' feature distributions. People use contrastive inferences to guide their referent choice, though size---and not color---adjectives prompt them to consistently choose the contrastive target over alternatives (Experiment 1). People also use color and size description to infer that a novel object is atypical of its category (Experiments 2 and 3): utterances like "the blue toma" prompt people to infer that tomas are *less* likely to be blue in general. However, these two inferences do not trade off substantially: people infer a described referent is atypical even when the descriptor was necessary to establish reference. We model these experiments in the Rational Speech Act (RSA) framework and find that it predicts both of these inferences. Overall, people are able to use contrastive inferences from description to resolve reference and make inferences about a novel object’s category, letting them learn more about new things than literal meaning alone allows.

authornote: |
  All data and code for analyses are available at https://github.com/-------/contrast.


keywords          : "concept learning; contrastive inference; word learning; pragmatics; communication; computational modeling"
wordcount         : "11021"
references        : ""

bibliography      : ["contrast.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"    
output            : papaja::apa6_pdf
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = '!tb', echo = FALSE, cache = TRUE, 
                      warning = FALSE, message = FALSE, 
                      sanitize = TRUE, fig.path='figs/', fig.width = 6,
                      fig.height = 3)
set.seed(42)
options(digits=3, dplyr.summarise.inform = FALSE)
```

```{r libraries, cache = FALSE}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(lme4)
library(broom)
library(broom.mixed)
library(here)
library(english)
library(ggthemes)
library(papaja)
library(gridExtra)
library(glue)
library(directlabels)
library(tidyboot)
library(lmerTest)
library(knitr)
library(rwebppl)
library(ggridges)
logit <- function(x) {log(x/(1-x))}
```

```{r set-theme, cache = FALSE}
theme_set(theme_few(base_size = 10) + theme(legend.position = "none"))
```

```{r make-text-vars}
make_text_vars <- function(df, term_name, term_filter = NULL) {
  if(!is.null(term_filter)) {
    filtered_df <- df %>%
      filter(term == term_filter) 
  } else{
    filtered_df <- df
  }
    
  walk(c("estimate", "statistic", "p.value"), 
      ~assign(glue("{term_name}_{.x}"), 
              filtered_df %>% pull(!!.x), 
         envir = globalenv()))
}
```

An utterance can say much more about the world than its literal interpretation might suggest. For instance, if you hear a colleague say "We should hire a female professor," you might infer something about the speaker's goals, the makeup of a department, or even the biases of a field—none of which is literally stated. These inferences depend on recognition that a speaker's intended meaning can differ from the literal meaning of their utterance, and the process of deriving this intended meaning is called *pragmatics*. Frameworks for understanding pragmatic inference posit that speakers tend to follow general principles of conversation—for instance, that they tend to be relevant, brief, and otherwise helpfully informative [@grice1975logic; @sperber1986relevance; @clark_pragmatics_1990]. When a speaker deviates from these principles, a listener can reason about the alternative utterances the speaker might have said and infer some intended meaning that goes beyond the literal meaning of their utterance.

Pragmatic inference is also a potentially powerful mechanism for learning language. People can learn the meanings of words by tracking statistical properties of their literal meaning alone [@yu2007], but reasoning about a speaker's intended meaning—and not just the words they say—may support more rapid and accurate learning [@frank2009]. For example, @akhtar_role_1996 showed that young children can infer the meaning of a new word by using the principle that people tend to remark on things that are new and interesting to them. In their study, an experimenter leaves the room and a new toy emerges in her absence; once she comes back, the toy is familiar to the child but not to the experimenter. When she uses a novel name, "gazzer," the child can infer that the word refers to the toy that is novel to the experimenter, and not other toys the experimenter had already seen. Much other work has emphasized that children can use pragmatic principles to infer a word's meaning, as can adults [@bohn2019pervasive; @bohn_how_2021; @bohn_predicting_2022; @gelman_implicit_1985; @baldwin1998word; @frank2014; @tomasello2000social].

One potential pragmatic tool for learning about referents is contrastive inference from description. To the extent that communicators strive to be minimal and informative, description should discriminate between the referent and some relevant contrasting set. This contrastive inference is fairly obvious from some types of description, such as some modifying adjunct phrases: "The door with the lock" clearly implies a contrasting door without one [@nietal]. The degree of contrast implied by more common descriptive forms, such as prenominal adjectives in English, is less clear: speakers do not always use prenominal adjectives minimally, often describing more than is needed to establish reference [@engelhardt_over-specified_2011; @mangold_informativeness_1988; @pechmann_incremental_1989]. Nevertheless, @sedivy_achieving_1999 showed that people can use these inferences to resolve referential ambiguity in familiar contexts. When asked to "Pick up the tall cup," people directed their attention more quickly to the target when a short cup was present, and did so in the period before they heard the word "cup." Because the speaker would not have needed to specify "tall" unless it was informative, listeners were able to use the adjective to direct their attention to a tall object with a shorter counterpart. Subsequent work using similar tasks has corroborated that people can use contrastive inferences to direct their attention among familiar referents and to infer which referents a speaker can see [@sedivy_pragmatic_2003; @aparicio_processing_2016; @ryskin2019information; @jara-ettinger_quantitative_2021].

But what if you didn't know the meaning of the key words in someone's utterance—could you use the same kind of contrastive inferences to learn about new words and categories? Suppose a friend asks you to "Pass the tall dax." Intuitively, your friend must have said the word "tall" for a reason. One possibility is that your friend wants to distinguish the dax they want from another dax they do not. In this case, you might look around the room for two similar things that vary in height, and hand the taller one to them. If, alternatively, you only see one object around whose name you don't know, you might draw a different inference: this dax might be a particularly tall dax. In this case, you might think your friend used the word "tall" for a different reason—not to distinguish the dax they want from other daxes around you, but to distinguish the dax they want from other daxes in the world. This would be consistent with data from production studies, in which people tend to describe atypical features more than they describe typical ones [@mitchell_2013; @westerbeek_2015; @rubio-fernandez_how_2016]. For instance, people almost always say "blue banana" to refer to a blue banana, but almost never say "yellow banana" to refer to a yellow one. 

In each of these cases, you would have used a pragmatic inference to learn something new. In the second case, you would have learned the name for a novel category "dax," and also something about the typical of size of daxes: most of them are shorter than the one you saw. In the first case, you would have resolved the referential ambiguity in the speaker's utterance. But would you have learned something about the typical size of daxes as well, beyond the daxes you observed? One possibility is that you would not: You can explain your friend's use of "tall" as being motivated by the need to distinguish between the two daxes in the room, and thus you should infer nothing about the other daxes in the world. If reference is the primary motivator of speakers' word choice, as implicitly assumed in much research [e.g., @pechmann_incremental_1989; @engelhardt_over-specified_2011; @arts_overspecification_2011], then people should draw no further inferences once the need for referential disambiguation can explain away a descriptor like "tall." On this *reference-first view*, establishing reference has priority in understanding the utterance, and any further inferences are blocked if the utterance is minimally informative with respect to reference. If, on the other hand, pragmatic reasoning weighs multiple goals simultaneously—here, reference and conveying typicality—people may integrate typicality as just one factor the speaker considers in using description. On this *probabilistic weighing view*, people can use description to make graded inferences about the referent's identity and about its category's features, and the fact that an adjective would have helped identify the referent does not completely block an inference about atypicality. 

In this paper, we present a series of experiments that test two ways in which people could use pragmatic inference to learn about novel categories. First, we examine whether listeners use contrastive inference to resolve referential ambiguity. In a reference game, participants saw groups of novel objects and were asked to pick one with a referring expression, e.g., "Find the small toma."  If people interpret description contrastively, they should infer that the description was necessary to identify the referent—that the small toma contrasts with some different-sized toma on the screen. We show that people can use contrastive inference—even with unfamiliar objects—to resolve reference and thus to learn the meaning of the new word "toma." 

Second, we test whether people use contrastive inference to learn about a novel category's feature distribution. Participants were presented with two interlocutors who exchange objects using referring expressions, such as "Pass me the blue toma." If people interpret description as contrasting with an object's category, they should infer that in general, few tomas are blue. Crucially, we vary the object contexts such that in some contexts, the adjective is necessary to establish reference, and in others, it is superfluous. Overall, we show that people can use contrastive inferences both to establish reference and to make inferences about novel categories' feature distributions, and that they do not trade off strongly between these two inferences. We extend a version of the Rational Speech Act model [@frank2014] that captures how listeners' reasoning about speakers reflects a graded integration of informativity with respect to both reference and typicality.

In order to determine whether people can use contrastive inferences to disambiguate referents and learn about categories' feature distributions, we use reference games with novel objects. Novel objects provide both a useful experimental tool and an especially interesting testing ground for contrastive inferences. These objects have unknown names and feature distributions, creating the ambiguity that is necessary to test referential disambiguation and category learning. Testing pragmatic inference in novel, ambiguous situations lays the groundwork to determine the role of pragmatic inference in learning language. Much work has focused on how pragmatic inference enriches literal meaning when the literal meaning is known—when the words and referents in play are familiar. Here, we ask: can people use pragmatic inferences from description to learn about unfamiliar things in the world? 

```{r child = "sections/experiment1.Rmd"}
```

```{r child = "sections/experiment2.Rmd"}
```

```{r child = "sections/experiment3.Rmd"}
```

# General Discussion

When we think about what someone is trying to communicate to us, we go far beyond the literal meanings of the words they say: we make pragmatic inferences about why they chose those particular words rather than other words they could have used instead. In most work on pragmatic reasoning, speakers and listeners share the same knowledge of language, and the question of interest is whether listeners can use their knowledge of language to learn something about the unknown state of the world. Here we focus on an even more challenging problem: Can pragmatic inference be used to learn about language and the world simultaneously?

In three studies we showed that people can use pragmatic inference to (1) learn the meaning of a novel word, (2) learn the typical features of the category described by this novel word, and (3) rationally integrate these two kinds of reasoning processes. In Experiment 1, we show that people can use descriptive contrast implied by adjectives like "big" or "blue" to resolve referential ambiguity to learn a new word; in the case of color, they shift substantially in the direction of the correct mapping, and in the case of size, they choose the correct mapping significantly more often than the incorrect one. In Experiments 2 and 3, we show that people infer that a noted feature is atypical of the object being referred to. Critically, people infer that the described feature is atypical even when the descriptor is helpful for referential disambiguation.

Why do people think that the mentioned feature is atypical even when its mention is helpful for referential disambiguation? If people use language for multiple goals—for example, both for reference and for description—then listeners should reason jointly about all of the possible reasons why speakers could have used a word. To determine what rational listeners would do in this circumstance, we developed an extension of the Rational Speech Act Framework that reasons both about reference and about the typical features of categories to which objects belong. The behavior of this model was closely aligned to the behavior we observed in people. Because rational inference is probabilistic rather than deterministic, the trade-off in the model is slight: descriptors still lead to atypicality inferences even when they are helpful for referential disambiguation. This work thus adds to the growing body of work extending the Rational Speech Act framework from reasoning about just reference to reasoning about other goals as well, such as inferring that speech is hyperbolic, inferring when speakers are being polite rather than truthful, and learning new words in ambiguous contexts [@goodman2016; @yoon2020; @kao2014; @frank2014; @bohn_how_2021; @bohn_predicting_2022].

In considering how people may integrate inferences about typicality and about reference, we raised two broad possibilities: (1) a *reference-first view*, whereby if an adjective was necessary for reference it would block an inference of atypicality completely, and (2) a *probabilistic weighing view*, whereby the goals of being informative with respect to reference and with respect to the category would trade off in a graded way. That is, we aimed to test whether there was a strong trade-off or a weak trade-off. People's behavior in our tasks is inconsistent with the reference-first view: inferences of atypicality were not blocked when an adjective was necessary for reference. On the other hand, our model implements the latter view and fits the data well, but we do not find significant evidence of a trade-off in our statistical tests of people's responses: the data are also compatible with there being no trade-off whatsoever. 

Because we find null effects of context, and our model predicts the effect of context to be small, we cannot tell from these experiments whether people make only slight trade-offs between these two communicative goals or only consider contrastive inferences with respect to typicality, without weighing it against reference. In the Supplemental Materials, we include an exploratory, combined analysis of the results from Experiments 2 and 3. In this combined analysis, we find that there is a slight effect of context, though no interaction between context and utterance type. We further provide a comparison of our model to a model that does not integrate referential context, and find that these models make similar predictions. Thus, people's behavior is broadly consistent with either rational integration of the goals of resolving reference and conveying typicality, or positing that they ignore referential context altogether when making inferences about typicality. Given prior work showing that communication about familiar concepts and objects is accurately modeled by weighing multiple communicative goals [@tessler_informational_2020; @kreiss2020production; @yoon2020], it is perhaps surprising that we did not find trade-offs between these two communicative goals with novel words and objects. Further work is necessary to tell whether effects of referential context are small or nonexistent, though we can rule out the position that there is an absolute trade-off between achieving reference and distinguishing an object from its category.

Our experiments use a particular kind of task context: alien fruits, spoken about by alien interlocutors. Would these effects generalize beyond these particular items, and this particular task? It is possible that people hold expectations about how the features of fruit are distributed—for instance, that they have stereotypical colors. These overhypotheses about how basic-level categories' features are distributed within a superordinate category [@kemp_learning_2007] may make people's inferences about fruit different from their inferences about other superordinate categories. In the Supplemental Materials we provide an additional demonstration that people make this inference about block shapes, which people likely do not expect to have stereotypical colors. However, it is an interesting and open question whether people's expectations about a category's feature distribution or their expectations about how often features of a category are mentioned would alter this effect. 

More broadly, people may make different kinds of inferences in more naturalistic communicative settings. In our task, people were asked to make several typicality judgments, which may have encouraged them to focus on how the aliens' utterances could help them learn about the world rather than focusing on other communicative goals such as reference. It is possible that people's inferences would reflect a clear tradeoff between reference and communicating atypicality if reference was a more salient communicative goal in the task. Further, it may be easier to attribute nuanced communicative goals to *people* talking about plausibly real things, rather than to alien characters. So, though we find people do use pragmatic inferences to resolve reference and learn about new categories in these artificial tasks, these inferences may play out differently in more naturalistic contexts with more communicative goals plausibly in play.

<!-- Participants were more successful in making pragmatic inferences about reference from size adjectives than from color adjectives. What makes size different from color? One possibility is that color adjectives are often used redundantly, and therefore receive less contrastive weight than adjectives consistently used to differentiate between referents. @sedivy_pragmatic_2003 puts forth such an account, finding that color adjectives tend not to be interpreted contrastively in eye-tracking measures except in contexts that make their use unlikely. In comparison, adjectives describing material (e.g., plastic) and size are interpreted contrastively, which corresponds to less redundant use of material and size adjectives in production. Further work is necessary to determine whether contrastive inferences hew to production norms, and whether implicit indications of contrast usually extend to explicit referent choice. -->

<!-- While the effects we show here are appropriately described as atypicality inferences from description, this result suggests that people's inferences about typicality are not simply inferring 'markedness' from the use of an adjective; any mention of an object can engender inferences about its category. -->

<!-- The relative robustness of contrastive inferences about typicality across contexts and adjective types compared to contrastive inferences among present referents raises questions about the relative importance of these two kinds of contrast in language understanding. Most prior work has focused on contrast with present referents as the main phenomenon of interest, with object typicality as a modulating factor; our results emphasize the role of contrast with an object’s category, particularly when ambiguity is at play. A reference-first view of utterance interpretation might predict that use of description would be largely explained away if the description was necessary for reference (e.g., the "red" in "red dax" is explained by a blue dax being present to distinguish from). Contrary to this possibility, we find that both our participants and a probabilistic model that integrates both referential utility and typicality make inferences of atypicality even when the adjective was necessary to establish reference. The model slightly weakens its inference of atypicality in this case, and participants' inferences do not significantly differ based on object context. Future work will explore whether people make subtle trade-offs between contrast with present referents and with the referent’s category. -->

Though the participants in our experiments were adults, the ability to disambiguate novel referents using contrast most obviously serves budding language learners---children. Contrastive use of adjectives is a pragmatic regularity in language that children could potentially exploit to establish word--referent mappings. Use of adjectives has been shown to allow children to make contrastive inferences among familiar present objects [@huangsnedeker2008; @davies_three-year-olds_2021]. When paired with other contrastive cues such as prosody, preschoolers can make inferences about novel object typicality [@horowitz_childrens_2016], and can use novel adjectives and nouns to restrict reference [@gelman_implicit_1985; @diesendruck_childrens_2006]. Future work should explore whether adjective contrast that is less scaffolded by other cues is a viable way for children to learn about novel concepts.  

<!-- In communicating with others, we have to refer to things in the world. But rather than being a neutral way to coordinate with interlocutors, even the barest referring expression conveys our beliefs about the referent in question and its relationship to other things.  -->

The core computation in pragmatic inference is reasoning about alternatives—things the speaker could have said and did not. Given that others are reasoning about these alternatives, no choice is neutral. In the studies in this paper, for instance, mentioning a feature led people to infer that the feature was less typical than if it had not been mentioned. But, conversely, *not* mentioning a feature led them to think that the feature was more typical than if they could not understand the meaning of the utterance at all—all communicative choices leak one's beliefs about the world. This has implications not only for learning about novel concrete objects, as people did here, but for learning about less directly accessible entities such as abstract concepts and social groups. These inferences can be framed positively, as ways for learners to extract additional knowledge that was not directly conveyed, but can also spread beliefs that the speaker does not intend. A core challenge is to understand how people reason about the many potential meanings a speaker might convey to *learn* about others' words for and beliefs about the world.
<!-- Contrastive inferences allow people to learn the meanings of new words and the typical features of new categories, pointing to a broader potential role of pragmatic inference in learning about the world.  -->

# Acknowledgements

<!--This research was funded by James S. McDonnell Foundation Scholar Award in Understanding Human Cognition \#220020506 to DY. The funding body had no involvement in the conceptualization, data collection, or analysis of this project.

The authors thank Ming Xiang and Susan Goldin-Meadow for guidance on early versions of this work and Benjamin Morris, Ashley Leung, Michael C. Frank, Ruthe Foushee, Judith Degen, and Robert Hawkins for feedback on the manuscript. Portions of this work were published in the proceedings of Experiments in Linguistic Meaning. The authors are grateful for feedback from reviewers and attendees of Experiments in Linguistic Meaning, the meeting of the Cognitive Science Society, and the Midwestern Cognitive Science Conference.-->

\newpage

# References

<div id = "refs"></div>
