---
title             : "Using contrastive inferences to learn about new words and categories"
shorttitle        : "Learning from contrastive inference"

author: 
  - name          : "Claire Bergey"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "5848 S. University Avenue, Chicago, IL 60637"
    email         : "cbergey@uchicago.edu"
  - name          : "Dan Yurovsky"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "The University of Chicago"
  - id            : "2"
    institution   : "Carnegie Mellon University"

 

abstract: |
  In the face of unfamiliar language or objects, description is one cue people can use to learn about both. Beyond narrowing potential referents to those that match a descriptor, listeners could infer that a described object is one that contrasts with other relevant objects of the same type (e.g., "The tall cup" contrasts with another, shorter cup). This contrast may be in relation to other present objects in the environment or to the referent’s category. In three experiments, we investigate whether listeners use descriptive contrast to learn new word-referent mappings and learn about novel categories' feature distributions. People use size adjectives contrastively to guide referent choice, though they do not do so using color adjectives (Experiment 1). People also use description to infer that a novel object is atypical of its category (Experiment 2). However, these two inferences do not trade off substantially: people infer a described referent is atypical even when the descriptor was necessary to establish reference. We model these experiments in the Rational Speech Act (RSA) framework and find it predicts both of these inferences. Overall, people are able to use descriptive contrast to resolve reference and make inferences about a novel object’s category, allowing them to learn more about new things than literal meaning alone allows.

authornote: |
  All data and code for these analyses are available at https://osf.io/3f8hy/?view_only=9a196db0444c4867bc899cc70a7a1e9c.


keywords          : "parent-child interaction; language development; communication"
wordcount         : "1385"
references        : "42"

bibliography      : ["contrast.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"    
output            : papaja::apa6_pdf
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = '!tb', echo = FALSE, cache = TRUE, 
                      warning = FALSE, message = FALSE, 
                      sanitize = TRUE, fig.path='figs/', fig.width = 6,
                      fig.height = 3)
set.seed(42)
options(digits=3, dplyr.summarise.inform = FALSE)
```

```{r libraries, cache = FALSE}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library(tidyverse)
library(lme4)
library(broom)
library(broom.mixed)
library(here)
library(english)
library(ggthemes)
library(papaja)
library(gridExtra)
library(glue)
library(directlabels)
library(tidyboot)
library(lmerTest)
library(knitr)
library(rwebppl)
library(ggridges)
logit <- function(x) {log(x/(1-x))}
```

```{r set-theme, cache = FALSE}
theme_set(theme_few(base_size = 10) + theme(legend.position = "none"))
```

```{r make-text-vars}
make_text_vars <- function(df, term_name, term_filter = NULL) {
  if(!is.null(term_filter)) {
    filtered_df <- df %>%
      filter(term == term_filter) 
  } else{
    filtered_df <- df
  }
    
  walk(c("estimate", "statistic", "p.value"), 
      ~assign(glue("{term_name}_{.x}"), 
              filtered_df %>% pull(!!.x), 
         envir = globalenv()))
}
```

An utterance can say much more about the world than its literal interpretation might suggest. For instance, the utterance "We should hire a female professor" may convey much about the speaker's goals, the makeup of a department, or even the biases of a field—none of which is literally stated. These pragmatic inferences are pervasive in everyday conversation: by reasoning about what someone says in relation to the context and what they might have said otherwise, we can glean more of their intended meaning. For example, \textcolor{red}{[some empirical thing with familiar referents akin to our example above]}.

But what if you didn't know the meaning of the key words in someone's utterance--could you use the same kind of pragmatic inferences to learn about new words and categories? Suppose a friend asked you to "Pass the tall dax." You might look around the room for two similar things that vary in height, and hand the taller one to them. Intuitively, your friend must have said the word "tall" for a reason [@grice1975logic]. One possibility is that your friend wanted to distinguish the dax they wanted from the dax they did not. People appear to make these kinds of inferences quite rapidly for objects they know; for instance, they already begin looking to a tall familiar object with a short competitor nearby--even if there are other tall objects around--as soon as they hear the word "tall" [@sedivy_achieving_1999]. 

If you only saw one object around whose name you didn't know, you might draw a different inference: this dax might be a particularly tall dax. In this case, you might think your friend used the word "tall" for a different reason--not to distinguish the dax they wanted and other daxes around you, but to distinguish the dax they want from other daxes in the world. This would be consistent with data from production studies, in which people tend to describe atypical features more than they describe typical ones [@mitchell_2013; @westerbeek_2015; @rubio-fernandez_how_2016]. For instance, people almost always say "blue banana" to refer to a blue banana, but almost never say "yellow banana" to refer to a yellow one. 

In each of these cases, you would have used a pragmatic inference to learn something new. In the second case, you would have learned the name for a novel category "dax," and also something about the typical of size of daxes: most of them are shorter than the one you saw. In the first case, you would have also learned a new word, but should have you learned something about the typical size of daxes as well, beyond the two daxes you observed? One possibility is that you should not: You can explain your friend's use of "tall" as being motivated by the need to distinguish between the two daxes in the room, and thus you should infer nothing about the other daxes in the world. If reference is the primary motivator of speakers' word choice, as implicitly assumed in much research \textcolor{red}{[cites]}, then people should draw no further inferences once the need for referential disambiguation can explain away a descriptor like "tall." 

If, on the other hand, pragmatic reasoning is probabilistic, people may integrate typicality as just one factor the speaker weighs in using description, leading to graded inferences about the referent's identity and about its category.

In this paper, we present a series of experiments to test whether and how listeners make inferences about novel referents using descriptive contrast. First, we examine whether listeners use descriptive contrast to resolve referential ambiguity. In a reference game, participants see groups of novel objects and are asked to pick one with a referring expression, e.g., "Find the blue toma." If participants interpret description contrastively, they should infer that the description was necessary to identify the referent--that the blue toma contrasts with some other-colored toma on the screen. Using this contrastive inference, they can resolve referential ambiguity, choosing a blue object with a similar non-blue counterpart rather than a blue object with no similar counterpart nearby. Second, we test whether listeners use descriptive contrast to make inferences about a novel object's category. Participants are presented with two interlocutors who exchange objects using referring expressions, such as "Pass me the blue toma." If participants interpret description as contrasting with an object's category, they should infer that in general, few tomas are blue. Crucially, we vary the object contexts such that in some contexts, the adjective is necessary to establish reference, and in others, it is superfluous. Overall, we show that people can use contrastive inferences both to establish reference and to make inferences about novel categories' feature distributions, and that they do not trade off strongly between these two inferences. We extend a version of the Rational Speech Act model to show that listeners' reasoning about speakers reflects a graded integration of informativity with respect to both reference and typicality.

In order to determine whether people can use prenominal adjective contrast to disambiguate referents, and how those inferences are affected by adjective type, we use reference games with novel objects. Novel objects provide both a useful experimental tool and an especially interesting testing ground for contrastive inferences. These objects have unknown names and feature distributions, creating the ambiguity that is necessary to test referential disambiguation and category learning. They have unknown names and feature distributions, creating the ambiguity necessary for our test of referential disambiguation. But the ability to disambiguate novel referents, or to establish reference with incomplete information, is also the broader problem of learning about the world. We know that distributional information in the world affects people’s pragmatic use and interpretation of description [@sedivy_pragmatic_2003-2, @westerbeek_2015]. Here, we ask: can people use pragmatic inferences from description to learn about unfamiliar things in the world?


```{r child = "sections/experiment1.Rmd"}
```

```{r child = "sections/experiment2.Rmd"}
```

```{r child = "sections/experiment3.Rmd"}
```

# General Discussion

When we think about what someone is trying to communicate to us, we go far beyond the literal meanings of the words they say. Instead, we make pragmatic inferences about why they chose those particular words rather than other words they could have used instead. This kind of reasoning allows us to draw scalar implicatures (e.g. "some" means "some but not all"), identify negative beliefs from the absence of positive language in recommendation letters, and to make the kind of typicality inferences we studied here. In most work on pragmatic reasoning, speakers and listeners share the same knowledge of language, and the question of interest is whether listeners can use their knowledge of language to learn something about the unknown state of the world. Here we focus on an even more challenging problem: Can pragmatic inference be used to learn about language and the world simultaneously?

In three studies we showed that people can use pragmatic inference to (1) learn the meaning of a novel word, (2) learn the typical features of the category described by this novel word, and (3) rationally integrate these two kinds of reasoning processes. In Experiment 1, we show that people can use descriptive contrast implied by adjectives like "big" or "blue" to resolve referential ambiguity to learn a new word. In Experiments 2 and 3, we show that people use the presence of the same kind of descriptor to infer that the noted feature is atypical of the object being referred to. Critically, people infer that the described feature is atypical even when the descriptor is helpful for referential disambiguation--although the size of the atypicality inference is slightly reduced.

Why do people think that the mentioned feature is atypical even when its mention is helpful for referential disambiguation? If people use language for multiple goals--for example both for reference and for description, then listeners should reason jointly about all of the possible why speakers could have used a word when they hear it. To determine what rational listeners would do in this circumstance, we developed an extension of the Rational Speech Act Framework that reasons both about reference and about the typical features of categories to which objects belong. The behavior of this model was closely aligned to the behavior we observed from people. Because rational inference is probabilistic rather than deterministic, descriptors still lead to atypicality inferences even when they are helpful for referential disambiguation. This work thus adds to the growing body of work extending the Rational Speech Act framework from reasoning about just reference to reasoning about other goals as well, such as inferring that speech is hyperbolic (e.g. waiting "a million years" means waiting a long time), inferring when speakers are being polite rather than truthful, and learning new words in ambiguous contexts [@goodman2016; @kao2014; @yoon2020; @frank2014].

In Experiment 1, participants notably failed to use color adjectives contrastively in choosing referents. What makes size different from color? One possibility is that color adjectives are often used redundantly, and therefore receive less contrastive weight than adjectives consistently used to differentiate between referents. Sedivy (2003) puts forth such an account, finding that color adjectives tend not to be interpreted contrastively in eye-tracking measures except in contexts that make their use unlikely. In comparison, adjectives describing material (e.g., plastic) and size are interpreted contrastively, which corresponds to less redundant use of material and size adjectives in production [@sedivy_pragmatic_2003-2]. Further work is necessary to determine whether contrastive inferences hew to production norms, and whether implicit indications of contrast usually extend to explicit referent choice.

While the effects we show here are appropriately described as atypicality inferences from description, this result suggests that people's inferences about typicality are not simply inferring 'markedness' from the use of an adjective; any mention of an object can engender inferences about its category.

The relative robustness of contrastive inferences about typicality across contexts and adjective types compared to contrastive inferences among present referents raises questions about the relative importance of these two kinds of contrast in language understanding. Most prior work has focused on contrast with present referents as the main phenomenon of interest, with object typicality as a modulating factor; our results emphasize the role of contrast with an object’s category, particularly when ambiguity is at play. A reference-first view of utterance interpretation might predict that use of description would be largely explained away if the description was necessary for reference (e.g., the 'red' in 'red dax' is explained by a blue dax being present to distinguish from). Contrary to this possibility, we find that both our participants and a probabilistic model that integrates both referential utility and typicality make inferences of atypicality even when the adjective was necessary to establish reference. The model slightly weakens its inference of atypicality in this case, and participants' inferences do not significantly differ based on object context. Future work will explore whether people make subtle trade-offs between contrast with present referents and with the referent’s category.

Though the participants in our experiments were adults, the ability to disambiguate novel referents using contrast most obviously serves budding language learners: children. Contrastive use of adjectives is a pragmatic regularity in language that children could potentially exploit to establish word--referent mappings. Further, use of adjectives has been shown to allow children to make contrastive inferences among familiar present objects [@huangsnedeker2008] and, when paired with other contrastive cues such as prosody, about novel object typicality [@horowitz_childrens_2016]; future work will explore whether adjective contrast alone is a viable learning tool in early childhood. Tasks using a mixture of novel adjectives and words suggest that children as young as 3 can make contrastive inferences about adjectives [@gelman_implicit_1985; @diesendruck_childrens_2006; @huangsnedeker2008]. Contrastive inferences allow people to learn the meanings of new words and the typical features of new categories, pointing to a broader potential role of pragmatic inference in learning about the world. 

# Conclusion

Taken together, these experiments show that people use contrastive inference to map novel words to novel referents and to make inferences about the typicality of novel referents' features. Hearing "small toma" allows people to narrow possible referents not only to small objects, but objects with larger counterparts nearby. Hearing "big toma" leads them to think that tomas are less likely to be that size. These two inferences do not trade off substantially: a referential use of description does not block an inference of atypicality. Rather, consistent with our model, people's typicality judgments are 

# Acknowledgements

This research was funded by a James S. McDonnell Foundation Scholar Award to the last author.

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
