# Experiment 3

```{r e3-read-data}
e3_data <- read_csv(here("data/exp3_turk_data.csv"))

# participants who have unbalanced numbers of trials in each condition
e3_exclude <- e3_data %>% count(subid, utttype, searchtype) %>% filter(n == 2 | n == 3)

e3_keep_subjs <- e3_data %>%
  filter(searchtype == "colorcheck", chosetarget == TRUE, attncheckscore >= 6) %>%
  filter(!(subid %in% e3_exclude$subid)) %>%
  group_by(subid) %>%
  count() %>%
  filter(n == 4)

e3_model_data <- e3_data %>%
  filter(subid %in% e3_keep_subjs$subid) %>%
  filter(trialtype != 0)

e3_subj_data <- e3_data %>%
  filter(subid %in% e3_keep_subjs$subid) %>%
  filter(trialtype != 0) %>%
  mutate(searchtype = if_else(searchtype == "differentshapes",
                                      "different", searchtype)) %>%
  mutate(rtsearch = rtsearch - 6500) %>% # time before selections can be made
  mutate(log_rt = log(rtsearch)) %>%
  mutate(adjective = if_else(utttype == "adj", "adjective noun", utttype),
         adjective = if_else(adjective == "noutt", "alien utterance", adjective),
         adjective = if_else(adjective == "noadj", "noun", adjective),
         adjective = factor(adjective, levels = c("noun", "adjective noun", "alien utterance")),
         searchtype = factor(searchtype, levels = c("contrast", "different")))

e3_mean_data <- e3_subj_data %>%
  group_by(searchtype, adjective, condition) %>%
  tidyboot_mean(percentage)

e3_model <- lmer(percentage ~ condition * adjective * searchtype +
                (adjective | subid),
              control = lmerControl(optimizer = "bobyqa"),
              data = e3_subj_data) %>%
  tidy() %>%
  filter(effect == "fixed")
```


In Experiments 1 and 2, we established that people can use contrastive inferences to resolve referential ambiguity and to make inferences about the feature distribution of a novel category. Additionally, in Experiment 2, we found that these two inferences do not seem to trade off substantially: even if an adjective is necessary to establish reference, people infer that it also marks atypicality. We also found that inferences of atypicality about color and size adjectives pattern very similarly, though their baseline typicality is shifted, while color and size are not equally contrastive with respect to referential disambiguation.


To strengthen our findings in a way that would allow us to better detect potential trade-offs between these two types of inference, here we replicate Experiment 2 in a larger sample of participants.  [ some explanation of why the new control condition is interesting as well ...]

## Method

### Participants.

Four hundred participants were recruited from Amazon Mechanical Turk. Half of the participants were assigned to a condition in which the critical feature was color (red, blue, purple, or green), and half of the participants were assigned to a condition in which the critical feature was size (small or big).

### Stimuli & Procedure.

The stimuli and procedure were identical to those of Experiment 2, with the following modifications. Two factors, utterance type and object context, were fully crossed within subjects. Object context had two levels: within-category contrast and between-category contrast. In the within-category context condition, Alien B possessed the target object and another object of the same shape, but with a different value of the critical feature (color or size). In the between-category contrast condition, Alien B possessed the target object and another object of a different shape, and with a different value of the critical feature. Thus, in the within-category contrast condition, the descriptor is necessary to distinguish the referent; in the between-category contrast condition it is unnecessary but potentially helpful. There were three utterance types: adjective, no adjective, and alien utterance. In the two alien utterance trials, the aliens spoke using completely unfamiliar utterances (e.g., "Zem, noba bi yix blicket"). Participants were told in the task instructions that sometimes the aliens would talk in a completely alien language, and sometimes their language will be partly translated into English. To keep participants from making inferences about the content of the alien utterances using the utterance content of other trials, both alien language trials were first; other than this constraint, trial order was random. We manipulated the critical feature type (color or size) between subjects.

After completing the study, participants were asked to select which of a set of alien words they had seen previously during the study. Four were words they had seen, and four were novel lure words. Participants were dropped from further analysis if they did not respond to at least 6 of these 8 correctly (above chance performance as indicated by a one-tailed binomial test at the $p = .05$ level). Additionally, six participants were excluded because their trial conditions were not balanced due to an error in the run of the experiment. This resulted in excluding XX participants, leaving XX for further analysis.  

```{r mean-data}
means <- e2_subj_data %>%
  group_by(condition, adjective, searchtype, subid) %>%
  gather(measure, value, percentage, log_rt) %>%
  group_by(condition, adjective, searchtype, measure, subid) %>%
  summarise(value = mean(value)) %>%
  tidyboot_mean(value)
```

```{r meanplotinference, fig.env = "figure", fig.width=8, fig.cap="The proportion of the novel category participants judged to have the feature of the target object, by condition. The left panel shows judgments on trials in which no adjective was used in the referring expression (e.g., 'Pass me the blicket'), and the right panel shows judgments on trials in which an adjective was used (e.g., 'Pass me the [purple/small] blicket'). This is crossed by the type of object context (contrast, different, same) on the x-axis."}
ggplot(filter(means, measure == "percentage"),
       aes(x = adjective, color = condition)) + 
  facet_wrap(~searchtype) + 
  geom_pointrange(aes(y = empirical_stat, ymin = ci_lower, ymax = ci_upper), 
                      position = position_dodge(.5)) +
  ylab("Prevalence judgment") +
  xlab("Context type") +
  theme(text = element_text(size=20)) +
  scale_color_ptol()
```



```{r e3-models}
# all prereg'd models below

e3_utt_model_no_alien <- e3_model_data %>%
  filter(utttype != "noutt") %>%
  mutate(utttype = factor(utttype, levels = c("noadj", "adj"))) %>%
  lmer(percentage ~ utttype + (utttype|subid), data = .) %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  mutate(p.value = printp(p.value))

make_text_vars(e3_utt_model_no_alien, "e3_adj_no_alien", "utttypeadj")

e3_utt_model <- e3_model_data %>%
  mutate(utttype = factor(utttype, levels = c("noutt", "noadj", "adj"))) %>%
  lmer(percentage ~ utttype + (utttype | subid), data = .) %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  mutate(p.value = printp(p.value))

walk2(c("e3_utt_model_adj", "e3_utt_model_noadj"), 
      c("utttypeadj", "utttypenoadj"), 
      ~ make_text_vars(e3_utt_model, .x, .y))

# model did not converge with maximal slopes
e3_full_model <- e3_model_data %>%
  mutate(utttype = factor(utttype, levels = c("noutt", "noadj", "adj"))) %>%
  lmer(percentage ~ utttype * searchtype * condition + 
         (utttype | subid), data = .) %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  select(-effect, -group) %>%
  mutate(p.value = papaja::printp(p.value))

walk2(c("e3_full_noadj", "e3_full_adj", "e3_full_size", "e3_full_diffshapes", "e3_full_adj_size"), 
      c("utttypenoadj", "utttypeadj", "conditionsize", "searchtypedifferentshapes", "utttypeadj:conditionsize"), 
      ~ make_text_vars(e3_full_model, .x, .y))

e3_full_model_no_alien <- e3_model_data %>%
  filter(utttype != "noutt") %>%
  mutate(utttype = factor(utttype, levels = c("noadj", "adj"))) %>%
  lmer(percentage ~ utttype * searchtype * condition + 
         (utttype | subid), data = .) %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  select(-effect, -group) %>%
  mutate(p.value = papaja::printp(p.value))

make_text_vars(e3_full_model_no_alien, "e3_noalien_adj", "utttypeadj")

```

```{r e3-rsa, eval = FALSE}
e3_null_inference <- webppl(program_file = 
                                         here("webppl/e3_null.wppl"))

e3_wppl_null_data <- e3_null_inference %>%
  pivot_wider(values_from = "value", names_from = "Parameter") %>%
  mutate(p = as.numeric(p)) %>%
  group_by(obj) %>%
  summarise(p = 100 * mean(p)) %>%
  filter(obj == "red toma") %>%
  rename(empirical_stat = p) %>%
  mutate(adjective = "alien utterance") %>%
  select(-obj) %>%
  expand_grid(condition = c("color", "size"),
              searchtype = c("different", "contrast"))
  
```

```{r extra-models}
no_baseline_model <- e3_model_data %>%
  filter(utttype != "noutt") %>%
  lmer(percentage ~ utttype + searchtype  + condition + 
         (utttype | subid), data = .) %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  select(-effect, -group) %>%
  mutate(p.value = papaja::printp(p.value))

size_only_model <- e3_model_data %>%
  filter(condition == "size") %>%
  mutate(utttype = factor(utttype, levels = c("noutt", "noadj", "adj"))) %>%
  lmer(percentage ~ searchtype * utttype +  
         (utttype | subid), data = .) %>%
  tidy() %>%
  filter(effect == "fixed") %>%
  select(-effect, -group) %>%
  mutate(p.value = papaja::printp(p.value))

```

```{r e3-wppl-plot, eval = FALSE}
e3_wppl_data <- e2_wppl_data %>%
  bind_rows(e3_wppl_null_data) %>%
  filter(searchtype != "same")

ggplot(e3_mean_data, aes(x = adjective, y = empirical_stat, 
                         color = searchtype, fill = searchtype)) + 
  facet_wrap(~ condition) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(.5)) + 
  theme(legend.position = "top") + 
  geom_col(data = e3_wppl_data, width = .5, 
           position = position_dodge(.5), alpha = .5) + 
  geom_hline(aes(yintercept = 50), linetype = "dashed")

```

## Results

We began by fitting a pre-registered maximum mixed-effects linear model: effects utterance type (alien utterance, adjective, or no adjective; alien utterance as reference level), context type (within category or between category), and critical feature (color or size) as well as all interactions and random slopes of utterance type and context type nested within subject. Random effects were removed until the model converged, which resulted in a model with all fixed effects, all interactions and a random slope of utterance type by subject. The final model revealed a significant effect of the no adjective utterance type compared to the alien utterance type ($\beta =$ `r e3_full_noadj_estimate`, $t =$ `r e3_full_noadj_statistic`, $p =$ `r e3_full_noadj_p.value`) and a marginal effect of the adjective utterance type compared to the alien utterance type ($\beta =$ `r e3_full_adj_estimate`, $t =$ `r e3_full_adj_statistic`, $p =$ `r e3_full_adj_p.value`). The effects of context type and adjective type were not significant ($\beta_{between} =$ `r e3_full_diffshapes_estimate`, $t_{between} =$ `r e3_full_diffshapes_statistic`, $p_{between} =$ `r e3_full_diffshapes_p.value`; $\beta_{size} =$ `r e3_full_size_estimate`, $t_{size} =$ `r e3_full_size_statistic`, $p_{size} =$ `r e3_full_size_p.value`), and there was a significant interaction between the adjective utterance type and the size condition ($\beta =$ `r e3_full_adj_size_estimate`, $t =$ `r e3_full_adj_size_statistic`, $p =$ `r e3_full_adj_size_p.value`). Thus, participants inferred that an object referred to in an intelligible utterance with no description was more typical of its category on the target feature than an object referred to with an alien utterance. They also inferred that an object referred to in an intelligible utterance with description was marginally less typical than an object referred to with an alien utterance, and this effect was slightly stronger in the size condition. They did not substantially adjust their inferences based on the object context.

Given that interpretation of these results with respect to the alien utterance condition can be difficult, we pre-registered a version of the same full model excluding alien utterance trials. This model revealed a significant effect of utterance type: participants' prevalence judgments were lower when an adjective was used than when it was not ($\beta =$ `r e3_noalien_adj_estimate`, $t =$ `r e3_noalien_adj_statistic`, $p =$ `r e3_noalien_adj_p.value`). No other effects were significant. This replicates the main effect of interest in Experiment 2: that when an adjective is used in referring to the object, participants infer that the described feature is less typical of that object's category than when the feature goes unmentioned. 



# Discussion

In Experiment 3, we replicated the main finding of interest in Experiment 2: when a novel object's feature is described, people infer that the feature is rarer of its category than when it goes unmentioned. Again, this effect was consistent across both size and color adjectives, and people did not substantially adjust this inference based on how necessary the description was to distinguish among potential referents. We also added an alien language condition, in which the entire referring expression was unintelligible to participants, to probe people's priors on feature typicality. We found that in the alien language condition, people judged features to be roughly between the adjective utterance and no adjective utterance conditions, and significantly different from the no adjective utterance condition. In the alien language condition, people's prevalence judgments were roughly around our model's prevalence judgments after observing the objects on each trial and before any inferences about the utterance.

The similarity of people's prevalence judgments on the alien language condition raises the question: is this effect driven by an atypicality inference in the adjective conditions, or a *typicality* effect when the feature is unmentioned? Our results suggest that it is a bit of both. When someone mentions an object without extra description, the listener can infer that its features are likely more typical than their prior; when they use description, they can infer that its features are likely less typical. Because using an extra word--an adjective--is generally not thought of as the default way to refer to something, this effect is still best described as a contrastive inference of atypicality when people use description. However, the fact that people infer high typicality when an object is referred to without description suggests that, in some sense, there is no neutral way to refer: people will make broader inferences about a category from even simple mentions of an object.